---
title: "Heron and Egret Project Season Summary Form"
output:
  word_document:
params:
  zyear: 2020
  zcode: 599
  zspp: "GREG"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)


```


```{r}
# just load packages not needed anywhere else
library(flextable)
library(officer)
library(knitr)

source("https://raw.githubusercontent.com/scottfjennings/rawhep_to_HEPDATA/main/code/rawhep_to_HEPDATA_utility_functions.R")

options(scipen = 999)

# some options for table column widths; probably won't need to change these
lead.col.width = c(0.5, 0.7, 1)
date.col.width = 0.6

```

```{r}
# step 1: input the year you are working with and read in the wrangled Survey123 data
# zyear = 2020
wrangled_raw <- readRDS(here(paste("data/wrangled/wrangled_raw", zyear, sep = "_")))



# step 2: check which season summary sheets have already been created
# readRDS(here(paste("season_summary_forms/sheet_creation_logs/sheet_creation_log_", zyear, sep = ""))) %>% view()



# step 3: input the colony code and species you want to generate a season summary sheet for

# zcode = 599
# zspp = "GREG"


# step 4: check if the species nested in the selected colony and year; if not, do not need to create season summary sheet.
# wrangled_raw$nests %>% filter(species == zspp, code == zcode) %>% summarize(max(total.nests))


# if this value is > 0, can click knit above and the season summary sheet will be generated
# if this value = 0, append the didn't nest list:
#if(wrangled_raw$nests %>% filter(species == zspp, code == zcode) %>% summarize(max(total.nests)) == 0) {
#  knit_exit()
#}

```



```{r}

# define  2 functions used only here

```


```{r}
# make a blank table to record screener initials

screeners <- data.frame(year = as.character(zyear),
                        code = zcode,
                        species = zspp,
                        screener.1 = "not screened",
                        screener.2 = "not screened",
                        screening.notes = "")

```
\fontsize{12}{22}
Screener initials:
```{r ft.align="left"}
flextable(screeners) %>% 
  width(j = 1:3, width = lead.col.width) %>% 
  width(j = 4:5, width = 1) %>% 
  width(j = 6, width = 3)
```


<br>
```{r}
# make colony/effort summary box
# Alcatraz spp counted on different days so effort needs to be stored in wrangled at species level. then need to deal with that here
if(zcode == 70 & zspp %in% c("BCNH", "SNEG")) {
 observers.effort <- wrangled_raw$observers.effort %>% 
  filter(code == zcode, 
         species == zspp) %>%  
  select(code, species, colony, everything())
 
} else {
observers.effort <- wrangled_raw$observers.effort %>% 
  filter(code == zcode) %>% 
  mutate(species = zspp) %>% 
  select(code, species, colony, everything())
}

```

```{r}
flextable(observers.effort) %>% 
  fit_to_width(max_width = 8.5) %>% 
  width(width = c(0.5, 1, 1, 1, 1, 1, 1))

```
 *indicates recording observer

<br>

```{r}


# total nests  
total.nests <- wrangled_raw$nests %>% 
  filter(code == zcode, species == zspp) %>% 
  select(species, date, multiple.survey.num, total.nests, complete.count) %>%  
  arrange(date) %>% 
  date2monthday() %>% 
  monthday_completecount() %>% 
  select(-multiple.survey.num, -complete.count) %>% 
  pivot_wider(values_from = total.nests, names_from = date) %>% 
  mutate(variable = "total.nests") %>% 
  select(variable, everything())

peak_active <- wrangled_raw$nests %>% 
  filter(code == zcode, species == zspp) %>% 
  select(species, date, multiple.survey.num, peak.active, complete.count)  %>% 
  arrange(date) %>% 
  date2monthday() %>% 
  monthday_completecount() %>% 
  select(-multiple.survey.num, -complete.count) %>% 
  pivot_wider(values_from = peak.active, names_from = date) %>% 
  mutate(variable = "peak.active") %>% 
  select(variable, everything())
  

obs_initials <- wrangled_raw$nests %>% 
  filter(code == zcode, species == zspp) %>% 
  select(species, date, multiple.survey.num, obs.initials, complete.count) %>%  
  arrange(date) %>% 
  date2monthday() %>% 
  monthday_completecount() %>% 
  select(-multiple.survey.num, -complete.count) %>% 
  pivot_wider(values_from = obs.initials, names_from = date) %>% 
  mutate(variable = "obs.initials") %>% 
  select(variable, everything())



nest_totals_df <- rbind(total.nests, peak_active, obs_initials) %>% 
  mutate(code = zcode) %>% 
  select(code, species, everything())
    
```
<br>
Total nests - note "code" is the colony specific code number  
"*" indicates "no" entered for complete count field
```{r}
nest_totals <- date_table_splitter(nest_totals_df)
```

```{r}
nest_totals$sub_table1 
```
<br>
```{r}
nest_totals$sub_table2 
```
<br>
```{r}
nest_totals$sub_table3
```
<br>
```{r}
nest_totals$sub_table4
```
<br>
```{r}
nest_totals$sub_table5
```



<br>
Number of nests at each stage
```{r}

if(zspp %in% c("GREG", "GBHE")) {
nest_stages_df <- wrangled_raw$stages %>% 
  filter(code == zcode, species == zspp) %>% 
  mutate(stage = paste("stage", stage, sep = ".")) %>% 
  arrange(date) %>%
  date2monthday() %>%  
  mutate(num.nests = as.character(num.nests)) %>% 
  select(-multiple.survey.num) %>% 
  pivot_wider(id_cols = stage, names_from = date, values_from = num.nests) %>% 
  mutate(code = zcode,
         species = zspp) %>% 
  rename(variable = stage)


rop <- wrangled_raw$stages %>% 
  filter(code == zcode, species == zspp) %>%  
  distinct(code, multiple.survey.num, date, species, which.rop) %>%
  arrange(date) %>%
  date2monthday() %>%  
  pivot_wider(id_cols = c(code, species), names_from = date, values_from = which.rop) %>% 
  mutate(code = zcode,
         species = zspp,
         variable = "which.rop")

nest_stages_df <- full_join(nest_stages_df, rop) %>% 
  select(code, species, everything())


}
```

```{r}
if(zspp %in% c("GREG", "GBHE")) {
nest_stages <- date_table_splitter(nest_stages_df)
} else {
print(paste("No nest stage data for", zspp))
}
```
<br>

```{r}
if(zspp %in% c("GREG", "GBHE")) {
nest_stages$sub_table1
}
```
<br>
```{r}
if(zspp %in% c("GREG", "GBHE")) {
nest_stages$sub_table2
}
```
<br>
```{r}
if(zspp %in% c("GREG", "GBHE")) {
nest_stages$sub_table3
}
```
<br>
```{r}
if(zspp %in% c("GREG", "GBHE")) {
nest_stages$sub_table4
}
```
<br>
```{r}
if(zspp %in% c("GREG", "GBHE")) {
nest_stages$sub_table5
}
```
<br>
<br>

Stage 4 brood size date selection.
```{r}
# stage 4 brood sizes



if(zspp %in% c("GREG", "GBHE")) {

brood_sizes <- wrangled_raw$brood.sizes %>% 
  filter(code == zcode, species == zspp)  %>% 
  group_by(date) %>% 
  mutate(max.num.nest = max(num.nests)) %>% 
  ungroup() %>% 
  filter(max.num.nest > 0)
  
if(nrow(brood_sizes) > 0) {
brood_sizes_wide <- brood_sizes %>% 
  mutate(brd = paste("brd.", brd, sep = "")) %>% 
  mutate(date = as.character(date),
         date = ifelse(multiple.survey.num > 1, paste(date, multiple.survey.num, sep = "."), date)) %>% 
  select(-multiple.survey.num) %>% 
  pivot_wider(values_from = num.nests, names_from = brd) %>% 
  select(code, species, date, everything(), -max.num.nest) %>% 
  arrange(date)
} else {
  brood_sizes_wide <- data.frame(code = zcode,
                                 species = zspp,
                                 date = NA,
                                 brd.size.date = NA,
                                 stage5.nests = NA,
                                 brd.1 = NA,
                                 brd.2 = NA,
                                 brd.3 = NA,
                                 brd.4 = NA,
                                 brd.5 = NA)
}
}
```
<br>

```{r}
if(zspp %in% c("GREG", "GBHE")) {
flextable(brood_sizes_wide) %>% 
  autofit() %>% 
    width(c(1, 6:ncol(brood_sizes_wide)), width = 0.5)
} else {
print(paste("No stage 4 brood sizes for", zspp))
}
```
note: only dates with at least 1 stage 4 nest are shown 

```{r}
# disturbance
disturbance <- wrangled_raw$disturbance %>% 
  filter(code == zcode, species == zspp) %>%  
  select(code, species, date, observed.inferred = obs.inf, description, type, result)

```
<br>
Disturbance to colony
```{r}
flextable(disturbance) %>% 
  autofit() %>% 
  width(5, width = 3) %>% 
  fit_to_width(max_width = 8.5)
```


```{r}
# predators
predators <- wrangled_raw$predators %>% 
  filter(code == zcode, species == zspp)

```
<br> 
Nest predatory species
```{r}
flextable(predators) %>% 
  autofit()

```

<br>
```{r}

notes <- wrangled_raw$notes %>% 
  filter(code == zcode, species == zspp) %>%  
  mutate(date = as.character(date),
         date = ifelse(multiple.survey.num > 1, paste(date, multiple.survey.num, sep = "."), date)) %>% 
  select(-multiple.survey.num) %>% 
  rbind(., data.frame(code = zcode,
                      date = NA,
                      species = zspp,
                      note.type = "for.HEPDATA",
                      notes = "")) 

```

<br> 
Notes
```{r}
flextable(notes) %>% 
  width(1, width = .5) %>% 
  width(2:4, width = 1) %>% 
  width(5, width = 4) %>% 
  valign(valign = "top")

```

<br>
Season Summary stat sheet - DO NOT EDIT

```{r}
seas_summary_stat_sheet <- data.frame(num_total_nest_tables = length(nest_totals[-which(sapply(nest_totals, is.null))]),
                                      num_nest_stage_tables = ifelse(exists("nest_stages"), 
                                                                            length(nest_stages[-which(sapply(nest_stages, is.null))]),
                                                                            0))


```

```{r}
flextable(seas_summary_stat_sheet) %>% 
  autofit()

```

